---
title: "file.system"
date: "Last rendered: `r Sys.time()`"
output: 
  html_document:
    theme: journal
    toc: true
    toc_depth : 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

# **google drive folder**

<br>

## **info/**

Information regarding the project and data, including admin. Includes action plans and literature reviews

- ### **invoicing/**
    - **Leave Sheet.gsheet:** googlesheet of booked holiday
    - **`invoice_id`-Bird Train Networks `purchase_order_id`:** copies of issued invoices.
- ### **Literature/**
  useful literature

## **inputs/**

- ### **data/**
  input data and metadata folder. Many input folders (`{rmacroRDM}`) are required by the functions of  `rmacroRDM`. Please do not alter the structure of these folders. For more details see [**rmacroRDM README**](http://annakrystalli.github.io/bird_trait_networks/rmacroRDM.nb.html). 
  
    - #### **analytical/**  
        - **`master.csv`** : master (long) analytical dataset. Output of `wkf = "rmacroRDM"` in `match.workflow.R` 
        - **`master wide.csv`** : master wide analytical dataset. Output of `wkf = "rmacroRDM"` in `match.workflow.R` 
    - #### **metadata/ {rmacroRDM folder}**
        - **`data_log.csv`** : metadata on raw datasets used to compile analytical
        - **`vnames.csv`** : metadata on raw data variable name matching to master
        - **`outliers.csv`** : table identifying outliers in duplicate `spp:var` entries
        - **`outlier_refs.csv`** : references supporting outlier identification
    - #### **pre/ {rmacroRDM folder}** :  
      contains `meta.var` folders, used to supply data to the `rmacroRDM` compilation functions. Dataset details can be found in `data_log.csv`.
        - ##### **csv/**
            - **Amniote_Database_Aug_2015.csv** : preprocessed amniote database **data** csv.
            - **Table.csv** : preprocessed original data csv.
        - ##### **n/**
            - **Amniote_Range_Count_Aug_2015.csv** : preprocessed amniote database **n** csv.
        - ##### **notes/**
        - ##### **observer/**
        - ##### **qc/**
        - ##### **ref/**
            - **Amniote_Database_References_Aug_2015.csv** : preprocessed amniote database **ref** csv. 
    - #### **post/ {rmacroRDM folder}** :  
      Folder where processed csvs are stored. Should mirror folders of contents of **pre/** folder.
        - ##### **csv/**
            - **Amniote_Database_Aug_2015.csv** : postprocessed amniote database **data** csv.
            - **Table.csv** : postprocessed original data csv.
        - ##### **n/**
            - **Amniote_Range_Count_Aug_2015.csv** : postprocessed amniote database **n** csv.
            - **Amniote_Database_Aug_2015_n_group.csv** : cross reference table confirming metadata column name assignment. Required when metavar information for variables is missing. Also used to assign metavar column names to multiple traits (e.g. if the same **n** referes to the sample size on which multiple trait data are based on.
        - ##### **notes/**
        - ##### **observer/**
        - ##### **qc/**
        - ##### **ref/**
            - **Amniote_Database_References_Aug_2015.csv** : postprocessed amniote database **ref** csv.     
    - #### **taxo/ {rmacroRDM folder}** : contains taxonomic metadata.
        - ##### **match data/** : 
          data relating to the taxonomic match workflow. Includes:
            - `mmatch` and `mmatched` match output `.csv`s used to complete manual matching (deprecated: compiled to syn.links sheet). 
            - match `m .RData` object for each dataset matched to the project `spp.list` (ie in our case `D1` and the `tree`)
        - ##### **synonyms/** : 
          various synonym lists from which `syn.links` were extracted.
        - **100spp.csv** : list of 100 species for subset analysis.
        - **`an.ID`species_ranks.RData** : Loaded during `Phylonetworker_setup.R` in `wkf = "phylonetworker"`. Contains: 
            - **`spp.rank`** : rank of sum of each species datapoint contribution across all possible trait pairs.
        - **`dcode`**_**taxo.csv** : taxonomic data extracted from matched dataset `dcode`
        - **syn.links.csv** : synonym links lookup table. 
        - **TD_`an.ID`.Rdata** : table of taxonomic distincteness assessment outputs of trait data subset taxonomic makeup. Loaded during `Phylonetworker_setup.R` in `wkf = "phylonetworker"`.
    - #### **tree/** : 
      contains files relating to the phylogenetic tree used in the analysis
        - **tree.RData** : tree used in analysis. Extracted from `Stage2_MayrAll_Hackett_set10_decisive.tre`.

## **outputs/**

- ### **data/**
    - #### **hrg/** : 
      outputs of experimentation with fiting consensus hierarchical networks using the `hrg` package.
    - #### **imputed_data/** : 
      outputs of missing data imputation (for data preparation for `mgm` analysis) using different subsets of the original data. Subselection of data uses the ranks of variable and species to the total data pool and is accomplished by setting the **proportion of top var/species** to be used through parameters **`v`** and **`s`** respectively. 
        - file naming: **`an.ID`spp-v`v`-s`s`.Rdata** 
    - #### **match objects/** : 
      m object resulting from processing and matching D1 data. {rmacroRDM}
    - #### **mgm/** : 
    mixed graphical models. filenaming **`an.ID`-`v`-`s`-`min.n`-`min.cat`**. {mgm}
    - #### **networks/** : 
      rnetcarto network outputs (as .Rdata as well as flat .csv). {phylonetworker}
        - Filenames: **`an.ID`_net_mn`min.n`_`log`**...
    - #### **phylocors/** : 
      outputs of phylogenetically corrected trait pair correlation analysis. {phylonetworker} 
        - Filenames: **`an.ID`_phylocor_mn`min.n`_`log`_`pair.type`.Rdata**.   
        -   All `pair.type` results complied into **`all_phylocor_mn50_log_allDT.csv`**.
    - **`master.rda`** : master object resulting from {rmacroRDM} workflow
    


- ### **reports/**
    - #### **progress reports/** : 
      Rendered (html) progress reports 01-06 and associated files
    - #### **Results/** : 
      Rendered (html) results reports
        - **Categorical_phylocors.nb.html** : Report on distribution of expected GKtau between pair of traits simulated from the phylogeny. 
        - **CorNetwork_plots.html** : early report ("nn" trait pairs only) on trait association network structure. Flat network plots.
        - **data_gap_eval.nb.html** : Report on data availability across traits. Include assessment of taxonomic distinctness of available trait data subsambles.
        - **hrg.html** : Demo of fitting a hierarchical graphical model.
        - **mgm_viz.html** : Visualisation of mixed graphical model (mgm) network
        - **Network_viz_`an.ID`.nb** : Vizualisation of phylonetworker networks
        
        
***

# github folder

<br>

The main folder contains the scripts to the primary workflows. The rest of the materials are organised in a system of folders. 

- **Phylonetworker.R** : `phylonetworker` workflow.
- **README** : repo README (landing page on online github repo).
- **configure_sys_ref.R** : script to intiate an rmacroRDM project and set up system reference files. Includes launching the **`sys_ref_configurator` app**. {rmacroRDM}
- **file_setup.R** : the first script to be modified. Sets project local folder pathways. {project_ui}
- **mgm.R** : `mgm` workflow.
- **project_ui.R** : script sourced at the beginning of any analysis to initialise the required environment. 
- **rmacroRDM_data.R** : `rmacro` workflow. Merging of the original with the amniote dataset.
- **rmacroRDM_data.R** : `rmacro` workflow. Synonym matching of `tree` to master `spp.list`.  
- ## **R/**
  folder containing scripts either sourced by primary workflows, or relating to secondary workflows.  
    
    - **PhyloNetworker_functions.R** : functions sourced for a number of workflows (see [**project_ui README**](http://annakrystalli.github.io/bird_trait_networks/project_ui.nb.html) for more details) to process data, determine trait associations and fit and vixsualise networks.
    - **PhyloNetworker_setup.R** : workflow script, sourced to prepare data and associated objects for analysis. `{project_ui}` 
    - **add amniote synonyms.R** : script to extract synonyms from supplied synonym list with `Amniote_Databsae`
    - **goodmankruskal.R** : GK  τ determination workflow for categorical variables.
    - **helper_functions.R** : project wide helper functions. `{project_ui}`
    - **impute_mgm.R** : script sourced during `mgm` wkf to impute missing data prior to mode fitting. `{mgm}`
    - **load_dependencies.R** : script to load dependencies. `{project_ui}`
    - **load_environment.R** : script to load wkf environment (functions and perform pre-processing) during `project_ui`.
    - **load_files.R** : loads `ref_system` files. `{project_ui}`
    - **load_global.R** : . `{project_ui}`
    - **mgm_dataprep.R** : sourced data preparation script `{mgm}`
    - **pkgs.R** : . `{project_ui}`
    - **rerun_phylosig.R** : script to re-extract lambda for traits using trait pair data subsamples. `{phylonetworker}`
    - **select_tree.R** : script to select tree. `{phylonetworker}`
    - **test_impute.R** : test to determine trait contribution to imputation error. Computationally expensive. `{mgm}`
    - **test_impute_functions.R** : functions required for `test_impute.R`. `{mgm}`

- ## **docs/** 
  folder to which **[gh-pages](https://pages.github.com/) is set (the project website)**. Contains **html** files comprising **public project documentation**. Use script `render.R` to render site.
  
    - **GoodmanKruskal.nb.html** : `goodmankruskal` workflow documentation.
    - **Hierarchical_Networks.html** : Theory and `R` tools behind Hierarchical network analysis.
    - **Network_Data_Availability_`an.ID`.nb.html** : Data availability across all trait pairs in `an.ID` analytical dataset.
    - **PhyloNetworker_README.nb.html** : `phylonetworker` workflow documentation.
    - **data_gap_eval.nb.html** : analysis of taxonomic bias across trait pair data subsets.
    - **file_system.html** : project file documentation.
    - **mgm.html** : `mgm` workflow documentation.
    - **mgm_viz.nb.html** : `mgm` network report & visualisation
    - **missing_data_thoughts.nb.html** : notes on approaches to deal with missing data.
    - **pglsPhyloCor_README.nb.html** : `pglsPhyloCor()` function documenation. Used in `phylonetworker` wkf.
    - **project_README.nb.html** : project README.
    - **project_ui.nb.html** : `project_ui.R` documentation. Script is sourced to initialise an analysis environment within the project. Environment specified through character strings `wkf` and `param`.
    - **pulling_repo.html** : documentation for making a pull request.
    - **rmacroRDM.nb.html** : rmacroRDM workflow documentation (needs updating)
- ## **man/**  
  Folder contains the **Rmd** files used to produced **html** documentation in `docs/` folder. Use **`render.R`** to render **Rmd**s to **.html**.
  
    - **GoodmanKruskal.Rmd** : `goodmankruskal` workflow documentation.
    - **Hierarchical_Networks.Rmd** : Theory and `R` tools behind Hierarchical network analysis.
    - **Network_Data_Availability_`an.ID`.Rmd** : Data availability across all trait pairs in `an.ID` analytical dataset.
    - **PhyloNetworker_README.Rmd** : `phylonetworker` workflow documentation.
    - **file_system.Rmd** : project file documentation.
    - **missing_data_thoughts.Rmd** : notes on approaches to deal with missing data.
    - **pglsPhyloCor_README.Rmd** : `pglsPhyloCor()` function documenation. Used in `phylonetworker` wkf.
    - **project_README.Rmd** : project README.
    - **project_ui.Rmd** : `project_ui.R` documentation. Script is sourced to initialise an analysis environment within the
    project. Environment specified through character strings `wkf` and `param`.
    - **pulling_repo.Rmd** : documentation for making a pull request.
    - **render.R** : documentation render workflow. Batch render and set theme.
    - **rmacroRDM.Rmd** : rmacroRDM workflow documentation (needs updating).
- ## **params/ {project_ui}**
    scripts to define workflow parameters. Sourced during `project_ui` by specifying name of file to source through object `"param"` Create new parametarisation by creating and specifying new parameter scripts scripts or modifying original files and re-running.
    
    - **goodmankruskal.R** : goodmankruskal parameters.
    - **mgm.R** : `mgm` parameters.
    - **min.n.test.R** : minimum n test workflow parameters.
    - **phylonetworker.R** : `phylonetworker` parameters.
    - **project_ui.R** : global `project_ui` parameters.
    - **rmacro.R** : rmacroRDM parameters.
    - **var_imput_test.R** : assessment of variable imputation error contribution.
- ## **process/ {rmacroRDM}**  
  folder containing custom scripts for processing individual rmacroRDM datasets stored in the **csv/** folder. They are sourced when `process_csv()` is called. Scripts are sourced either:
    + **pre-processing** : any computation will be performed on the raw data.fame as first loaded from the **pre/** `rmacroRDM` folder. Processing script names **should be appended with `_pre`**.
    + **post-processing** : any computation will be performed on the data.fame produced post processing, prior to being saved in the **post/** `rmacroRDM` folder. 
    Processing script are **named as the dataset `file.name` and appended with `_post`**.
    
      - **Amniote_Database_Aug_2015_post.R** 
      - **Amniote_Database_Aug_2015_pre.R** 
      - **Amniote_Database_References_Aug_2015_pre.R** 
      - **Amniote_Range_Count_Aug_2015_pre.R** 
      - **Table_post.R** 
      - **Table_pre.R** 
- ## **reports/** 
    - ### **results/** 
      Folder contains **Rmd** documents used to generate results **html** documents in **`output.folder`/reports/results/**.
      
         - **Categorical_phylocors.Rmd** : Report on distribution of expected GKtau between pair of traits simulated from the phylogeny. 
        - **CorNetwork_plots.Rmd** : early report ("nn" trait pairs only) on trait association network structure. Flat network plots.
        - **Network_viz.Rmd** : Vizualisation of phylonetworker networks
        - **data_gap_eval.Rmd** : Report on data availability across traits. Include assessment of taxonomic distinctness of available trait data subsambles.
        - **hrg.Rmd** : Demo of fitting a hierarchical graphical model.
        - **mgm_viz.Rmd** : Visualisation of mixed graphical model (mgm) network


***

# **parameters**

<br>

Parameters for analysis environments are supplied through scripts in the **params/** folder and called by specifying a **`param`** character string of the name of the required **params/** file, e.g.
```{r}
param <- "phylonetworker.R"
```
prior to sourcing `project_ui.R`. The **`params/project_ui.R`** is a required global parameter file and is called as part of `project_ui` sourcing. Please do not delete any files. Add new ones to parametarise your own analysis.

## `project_ui` parameters:

- **na.strings** : character vector defining values to be encoded as `NA` when reading **csv**s.
- **fileEncoding** : character vector specifying global `fileEncoding` to be used when reading in csvs in primary workflows. rmacroRDM handles fileEncodings on an individual file basis and is not affected by this setting. But loadiing the master outputs of the `rmacroRDM` workflow in downstream analysis might required fiddling with this setting (especially for `metadata.csv` and `master.csv`). If you're getting funny characters on check out the [**fileEncoding_checker**](https://annakrystalli.shinyapps.io/fileEncoding_checker/) to test fileEncoding on your system.  

## `wkf` parameters:
- **an.ID** : character string identifying analysis. Either **`"all"`** to use all available species data or **`"100spp"`**, for subsetting data to the 100 species subset defined in **taxo/100spp.csv**.
- **min.n** : minimum number of shared datapoints required for a trait pair to be included in association analysis.
- **log** : logical, whether to log numeric traits, specified as `log.vars` through the logical **log** column in `metadata.csv`.
- **cutoff** : cut-off level used to delimit association between trait pairs. Defaults to 0.35
- **min.cat** : Minimum accepted count per category. if the count of any category falls below this threshold, categories affected are attempted to be re-encoded as a single "other" category. If this still doesn't result in > min.cat counts, the affected category and datpoints are removed. Defaults to 5.
- **remove_dtips** : whether to remove tips from tree which have resulted as duplicates through matching the taxonomy to the master spp.list. Defaults to `T` for smooth matching of tree to data.
**dist.types** : character vector. Either **`"all"`** to incidate inclusion of all distribution (data) types or a vector indicating the **`"mgm_types"`** to be included.
- **save** : T

### mgm parameters
- **v.p** : proportion of variables to be included in `mgm` analysis. Variables are selected from a lists of variables ranked according to their cumulative datapoint contribution across all trait pairs.
- **s.p** : proportion of species to be included in `mgm` analysis. Species are selected from a lists of variables ranked according to their cumulative datapoint contribution across all trait pairs.