---
title: "wkf: GoodmanKruskal"
date: "Last rendered: `r format(Sys.time(), '%d %b %Y %H:%M:%S')`"
output: 
  html_document:
    theme: paper

---
Demonstration of use of the GoodmanKruskal workflow. The data processing is based on `phylonetworker_setup.R` but includes a data processing stage from an `mgm` workflow.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning=F, cache=F)
```

## parametarise and launch analysis environment:

```{r, echo=TRUE}
wkf = "goodmankruskal"
param = "goodmankruskal.R"
file_setup_path <- "~/Documents/workflows/bird_trait_networks/file_setup.R"
source(file_setup_path)
source("~/Documents/workflows/bird_trait_networks/project_ui.R")

```

### analysis parameters 

```{r}
read_chunk(paste0(script.folder,"params/", param))
```

```{r pm-params, echo=T, eval=FALSE}

```


Here we select all categorical traits and species, by setting `v.p` (proportion of variables) to `r v.p`and `s.p` is set to `r s.p`.

```{r}
vars <- ms_vars[1:(as.integer(length(ms_vars)*v.p))]
spps <- spp.ranks$spp[1:(as.integer(nrow(spp.ranks)*s.p))]
spp <- data$species %in% spps

rownames(data) <- data$species

```


###create g object 

collect all information required to process and fit data (for consistency and processing with other workflows) and prepare data.

```{r, echo = T}

g <- list(data = data[spp, vars], 
          meta = metadata, 
          mgm_types = mgm_types, meta_types = meta_types, 
          log.vars = log.vars, 
          spp.list = data.frame(species = data[spp, c("species")]),
          v.p = v.p, s.p = s.p, phylo.match = phylo.match[spp,],
          tree = drop.tip(tree, setdiff(tree$tip.label, data$species[spp])))

source(paste(script.folder, "mgm_dataprep.R", sep = ""))

```

<br>

### subset to only categorical variables. 

```{r, echo = T}

gdk_data <- g$data[,names(mgm_types[mgm_types %in% c("c","b")])]
```
#### $n_{species}$: `r length(spp)`

#### $n_{var}$: `r ncol(gdk_data)`

`r names(gdk_data)`

#### ${min}_{{pair}_n}$: `r min.n`

<br>

### Calculate goodmankruskal $τ$ matrix.

```{r, echo=T}
gtau <- GKtauDataframe(gdk_data)
```



<br>

### **plot goodmankruskal $τ$ matrix**

```{r}
diag(gtau) <- NA

m.lab <- 120
m = list(
  l = m.lab,
  r = 50,
  b = m.lab,
  t = 50,
  pad = 4
)

plot_ly(z= gtau, x = colnames(gtau), y = rownames(gtau), type = plot.type) %>%
  layout(title="GKtau among categorical variables", margins = m, 
         yaxis = list(title = "", showgrid = F), 
         xaxis = list(title = "", showgrid = F))


```

<br>

***

### fit netcarto

```{r}
diag(gtau) <- 0
gtau[gtau < cutoff] <- 0

g <- graph.adjacency(gtau, mode = c("directed"), weighted = T)
e <- get.edgelist(g) %>% data.frame() %>% as.list()
e$w <- E(g)$weight

net <- rnetcarto::netcarto(e)

```

### plot network based on goodmankruskal $τ$ matrix

```{r}
v <- data.frame(var = factor(rownames(gtau), 
                             levels = rownames(gtau)),
                net[[1]][match(rownames(gtau), net[[1]]$name),])

e <- graph.adjacency(gtau, mode = c("directed"), weighted = T)  %>% get.edgelist(names = T) %>% 
  as.data.frame() %>% setNames(c("source", "target"))
e$source.index = match(e$source, v$name)-1
e$target.index = match(e$target, v$name)-1


w <- melt(as.data.frame(as.matrix(get.adjacency(g, attr="weight"))))
w <- w[w$value > 0,]
e <- data.frame(e, value = w$value)



n <- forceNetwork(Links = e, Nodes = v, Source = "source.index",
             Target = "target.index", Value = "value", NodeID = "var",
             Group = "module", opacity = 0.75, fontSize = 18, zoom = T)

n
```


#### plot flat bi-directional network

```{r}
plot(g, edge.arrow.size=.2, edge.curved=.1, vertex.size = 3, vertex.label=NA)
```

