---
title: "PhyloNetworker_README"
date: "Last rendered: `r format(Sys.time(), '%d %b %Y %H:%M:%S')`"
output: html_document
params:
  file_setup_path: ../file_setup.R
  save: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
attach(params)
source(file_setup_path)
```

```{r read_chunks, echo=FALSE}
read_chunk(paste0(script.folder, "PhyloNetworker.R"))
read_chunk(paste0(script.folder, "params/phylonetworker.R"))
read_chunk(paste0(script.folder, "R/Phylonetworker_setup.R"))
source(paste0(script.folder, "R/Phylonetworker_functions.R"))

```

***

<br>

# **setup**

First clear the workflow and set up the file system by **providing the path to `file_setup.R`**. Make sure `file_setup.R` also contains valid paths to the project files.


```{r pn-clear-wkf, eval=FALSE}

```
```{r pn-setup, message=FALSE, warning=FALSE}

```

<br>

## Call project_ui

To parametarise the analysis, `wkf` is set to **`"PhyloNetworker.R"`** and `param` to **`"phylonetworker.R"`**.
```{r pn-project-ui, warning=FALSE}

```

Next initialise the analysis by calling **`project_ui`**. This script calls a number of ***global*** and ***analysis specific*** setup scripts (for details see [**`project_ui_README`**](http://annakrystalli.github.io/bird_trait_networks/project_ui.nb.html)), including:

- **`params/phylonetworker.R`** parameter script
```{r pm-params}

```

- [**`PhyloNetworker_setup.R`**](https://github.com/annakrystalli/bird_trait_networks/blob/master/R/PhyloNetworker_setup.R)
- [**`PhyloNetworker_functions.R`**](https://github.com/annakrystalli/bird_trait_networks/blob/master/R/PhyloNetworker_functions.R)

*** 

### **`PhyloNetworker_setup.R`**

You don't need to worry about running this, calling `project_ui` does. But I'm just running through it so you understand how the data and objects used in the analysis have been processed and created.

#### First, files required for the analysis are loaded or created:
```{r pns-load-files, eval=FALSE}

```

These include:

- **`wide`:** trait data in wide format (output of `match_workflow.R`)
- **`spp100`:** list of 100 species names for subset analysis
- **`tree`:** the phylogenetic tree 
- **`phylo.match`** taxonomic reference table linking species in tree to the data.
- **`pro.pars`:** processing parameters for combining mean and individual sex data.
- **`alco.vars:`** (created) vector of variables to be allometrically corrected.

#### Prepare taxonomic & phylogenetic objects
```{r pns-remove-dup-dftips, eval=FALSE}

```

- remove species in wide data that correnspond to duplicate tree tip synonyms. This is to ensure smooth alignment of the tree with the trait data.
- trim tree to species list in de-duplicated data.

#### Merge mean and sexed traits and remove excluded traits

Use parameters in `proc.pars` which specifies which traits to be used as primary and which to be used to supplement to merge into single trait. Also create `.ssd` (sex dimorphism traits) from sexed traits using equation: `trait`.ssd $= log(M/F)$
```{r pns-sex-data-merge, eval=FALSE}

```

`sex_data_merge function`
```{r}
sex_data_merge
```

Use parameters in `proc.pars` to exclude indicated traits
```{r pns-var.delete, eval=FALSE}

```

#### Rename tree tip labels and create `spp.list`
Next, tree tip labels are renamed to match data taxonomy
```{r pns-rename-tip.labels, eval=FALSE}

```

and `spp.list` is created
```{r pns-create-spp.list, eval=FALSE}

```



#### Subset data (for an.ID = `100spp`)
```{r pns-subset-data, eval=FALSE}

```

#### Create analysis metadata

Subsets metadata to traits included in analysis. Creates metadata rows for merged and `.ssd` traits.
```{r pns-create-analysis-metadata, eval=FALSE}

```



#### `log.vars` are extacted from `an_meta`

```{r pns-get-log.vars, eval=FALSE}

```


#### Create data type reference objects
```{r pns-create-ref.obj, eval=FALSE}

```

- **`vg`:** grid of available trait (var) pairs and the number of datapoints avalable for each. Pairs are only included in pair $n$ greater than $n_{min}$. 
```{r}
head(vg)
```
- **`ms_vars`:** character vector containing all unique trait names available for analysis.

#### Trim `data` and `an_meta` to `ms_vars` traits
```{r pns-trim-data, eval=FALSE}

```

#### Correct categories in factor traits which are not integers
```{r pns-correct-cat, eval=TRUE}

```

#### create reference objects
```{r pns-types-ref}

```

**`vtypes`:** reference table linking `metadata` data.types to mgm data types

 - **`"g"`:** gaussian
 - **`"p"`:** poisson
 - **`"c"`:** categorical (including binomial)
 
 I've kept the mgm data type system because it reminds us to think of distributions and also treat binary and categorical as a single category, as required by the `phylonetworker` workflow.

```{r print.types}
vtypes

```

**`meta_types`:** named vector of trait metadata category types, matched to and named by `ms_vars`. 

```{r}
head(meta_types)
```

**`mgm_types`:** corresponding named vector of trait mgm category types, matched to and named by `ms_vars`. 
```{r}
head(mgm_types)
```

#### Center numeric data

```{r pns-center_dt}

```

#### Create master analysis reference objects
```{r pns-create-ref.obj2, eval=FALSE}

```

- **`ms_spp`:** character vector of species included in analysis
```{r}
head(ms_spp)
```
- **`spp.ranks`:** data.frame containing the total number of data points from all possible combinations of species datasets across all traits when this species is included in the analysis. Sorted in descending order of cumulative data availibility.

```{r}
head(spp.ranks)
```


***

<br>

# **Edge determinations**


#### load / get trait taxonomic diversity profiles.

This chunck calculates the taxonomic diversity of subsamples dictated by data availability by trait. For each trait, the expected taxonic distinctness from a random subsample of the full pool of species of size $n_{trait}$

```{r pn-get-TD, eval=TRUE}

```

```{r}
kable(head(format(TDdf, digits = 2)))
```


### Extract or load allometrically corrected residuals. 
(Although created, currently this isn't used in the analysis. On hold to further notice)
```{r pn-alco-correct, eval=FALSE}

```


### Create `vg_dt`
Dataframe of **mgm data types**, coresponding to trait pair columns of `vg`.  
```{r pn-get-vg_dt}

```

```{r}
head(vg_dt)
```


Next edge weights between traits are calculated by calling the function **`pglsPhyloCor`** for each **`pair`** of each **`pair.type`**.

#### pglsPhyloCor function

The phylocors function calculates **phylogenetically corrected correlations** between pairs of traits (defined by the *character vector of the trait names* supplied to argument `pair`), which are then used as edge weights in the network. The method is determined by the argument `pair.types`. 

  - `"nn"`: numeric vs numeric (includes poisson). If `log = T`, data for traits defined in `log.vars` are logged. Method uses `pgls()`.
  - `"nc"`: numeric vs categorical. (variables must be provided in the order "nc" in `data`). Method uses `pgls()`
  - `"cc"`: categorical vs categorical. Method combines phylogenetic trait coevolution simulations and `GoodmanKruskal()`.
  
  for more details see [**`pglsPhyloCor` README**](http://annakrystalli.github.io/bird_trait_networks/pglsPhyloCor_README.nb.html)
```{r}

```


###  numeric vs numeric (`pair.type = "nn"`)
```{r pn-nn, eval=FALSE}

```

###  numeric vs categorical (`pair.type = "nc"`)
```{r pn-nc, eval=FALSE}

```

###  categorical vs categorical (`pair.type = "cc"`) 

Edge determination for `pair.type = "cc"` proceeds in two steps

1. `pglsPhyloCors` returns a list (in order to retain additional pertinent information, such as the evolutionary model selected for trait simulation in the determination of phylogenetic GoodmanKruskal τ)

```{r pn-cc.list, eval=FALSE}

```

2. Information is extracted from the list and returned as a data.frame, alignable with `pglsPhyloCor` ouputs for `pair.type = "nn"` or `"nc"` (more details in [**`pglsPhyloCor` README**](http://annakrystalli.github.io/bird_trait_networks/pglsPhyloCor_README.nb.html).    
```{r pn-cc, eval=FALSE}

```

### Combine all pair.type results

```{r pn-comp.res, eval=FALSE}

```


```{r pn-read.type-res, echo=FALSE}

```

```{r}
kable(head(res), row.names = F, digits =2)
```

***

<br>

# **Network analysis**

Perform network analysis. The workflow uses the function `rnetcarto::netcarto()`. See further details on the [package](https://cran.r-project.org/web/packages/rnetcarto/vignettes/getting-started.html) and [theory](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2151742/). 

```{r pn-run-netcarto, eval=FALSE}

```

The results are saved as the output of the `rnetcarto()` function and also collapsed into a single data.frame, appended with some useful trait level details, and saved as a `.csv` on folder **outputs/data/networks/. A `...net-stats.csv` containing network level statistics is also saved.

Here's an example of the data.frame output. See **`NetworkViz.html`** for details or resulting data. 

```{r}
kable(head(run_netcarto(res, edge_det = "phylocor", alco, save = F)), 
      row.names = F, digits =2)
```

## Visualise network

To visualise the network analysis, render the `.Rmd` report [`"reports/results/Network_viz.Rmd""`](https://github.com/annakrystalli/bird_trait_networks/blob/master/reports/results/Network_viz.Rmd). 

Four reports are produced for each `an.ID` (total 8). A combination of each of determining network edges on the basis of either `"phylocor"` or `"p-value"`, and of displaying the resulting networks with nodes coloured according to either `"cat"` or `"module"`. Results reports are in the googledrive folder: **outputs/Reports/Results/network_visualisations/**.

```{r, eval=FALSE}
require(rmarkdown)
  file <- "Network_viz.Rmd"
  for(an.ID in c("all", "100spp")){
    for(edge_det in c("p-value", "phylocor")){
      for(node_colour in c("module", "cat")){
        render(paste("reports/results/", file, sep =""), 
               output_format = "html_notebook",
               #output_options = list(theme = theme),
               output_file = paste(output.folder,"Reports/Results/network_visualisations/",
                                   gsub(".Rmd", paste0("_",an.ID,
                                                       "_net_mn", min.n, 
                                                       "_", edge_det,
                                                       "_node-col-", node_colour,
                                                       if(alco){"_alco_"},
                                                       if(log){"_log"}), 
                                        file), ".nb.html", sep =""),
               params = list(wkf = "phylonetworker",
                             param = "phylonetworker.R",
                             file_setup_path = "~/Documents/workflows/bird_trait_networks/file_setup.R",
                             an.ID = an.ID, edge_det = edge_det,
                             node_colour = node_colour))
      }
    }
  }
```



