---
title: "Network data availability"
author: "Anna Krystalli"
date: "14 December 2015"
output: html_document
---


```{r, echo=F, message=FALSE, warning= F}

codewd <- "~/Documents/worflows/bird_trait_networks/"
datawd <- "~/"

output.folder <- "/Users/Anna/Google Drive/bird trait networks/outputs/"
input.folder <- "/Users/Anna/Google Drive/bird trait networks/inputs/data/"


require(dplyr)
require(plotly)
require(knitr)
require(RColorBrewer)

wide <- read.csv(file = paste(input.folder, "csv/","master wide.csv", sep = "")
               ,fileEncoding = "mac")

  spp100 <- unlist(read.csv(file = paste(input.folder,"csv/100spp.csv", sep= "")))



numerise <- function(x){if(all(grepl('^[0-9.]+$', x))) as.numeric(x) else x}

makeTransparent<-function(someColor, alpha=100)
{
  newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                              blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
}

calcTraitPairN <- function(data){
  
  vars <- names(data)[!names(data) %in% c("species", "synonyms")]
  
  var.grid <- expand.grid(vars, vars, stringsAsFactors = F)
  var.grid <- var.grid[var.grid[,1] != var.grid[,2],]
  
  indx <- !duplicated(t(apply(var.grid, 1, sort))) # finds non - duplicates in sorted rows
  var.grid <- var.grid[indx, ]
  
  countN <- function(x, data){sum(complete.cases(data[,c(x[1], x[2])]))}
  
  var.grid <- data.frame(var.grid, n = apply(var.grid, 1, FUN = countN, data = data))
  
}

#source("~/Documents/worflows/bird_trait_networks/Setup.R")

metadata <- read.csv(paste(input.folder, "metadata/","metadata.csv", sep = ""), 
                     stringsAsFactors = F, fileEncoding = "mac") %>% 
  apply(2, FUN = trimws) %>% data.frame(stringsAsFactors = F)


vnames <- read.csv(paste(input.folder, "metadata/","vnames.csv", sep = ""), 
                   stringsAsFactors = F, fileEncoding = "mac") %>% 
  apply(2, FUN = trimws) %>% data.frame(stringsAsFactors = F)
vnames[vnames == ""] <- NA

#__________________________________________________________________
### Data availability across **`D0`** bivariate combinations
```

Focusing on the 100 species list supplied, I've calculated counts of data available for each possible bivariate combination of variables in the dataset. Below is the distribution of counts across all bivariate combination.



```{r, echo=F, message=FALSE, warning= F}

wide <- wide[wide$species %in% spp100,]

# make grid with unique combinations of variables
var.grid <- calcTraitPairN(wide[,-(1:3)])
 vars <- unique(c(var.grid$Var1, var.grid$Var2))
  var.grid$np <- var.grid$n/dim(wide)[1]

# make matrix
m <- matrix(NA, ncol = length(vars), nrow = length(vars),
dimnames = list(vars, vars))
m[cbind(var.grid[,1], var.grid[,2])] <- var.grid$np
m[cbind(var.grid[,2], var.grid[,1])] <- var.grid$np

m[cbind(vars, vars)] <- 1

# order by categories
cat <- metadata$cat[match(dimnames(m)[[1]], metadata$master.vname)]

 cum.cor <- c() 
  for(var in dimnames(m)[[1]]){
   cum.cor <- c(cum.cor, mean(var.grid$np[var.grid$Var1 == var | var.grid$Var2 == var], na.rm = T))
  }

m.order <- order(cum.cor, decreasing = T)
labels <- paste(substr(cat, 1,2), dimnames(m)[[1]])[m.order]
m <- m[rev(m.order),m.order]

plot_ly(x = var.grid$n, type = "histogram") %>%
  layout(title = "Distribution of bivariate datapoint availability",
         xaxis = list(title = "number of datapoints available per bivariate combination"),
         yaxis = list(title = "frequency"))


```


I've also plotted interactively (*ie hover over the matrix for details of each cell*) the matrix of bivariate counts, sorted by mean variable data availability when considered across all possible variable combinations
  
  
```{r, echo=F, message=FALSE, warning= F}
col <- brewer.pal(5, "YlGnBu")

plot_ly(z = m, x = labels, y = rev(labels), type = "heatmap", 
        #color = col, 
        name = "Bivariate proportional datapoint availability",
        colors = col,
        hoverinfo = "z+x+y", text = text) %>%
  
  layout(margin = list(l = 150,
                       r = 80,
                       b = 120,
                       t = 60),
         title = "Matrix of proportionate bivariate datapoint availability",
         xaxis = list(title = "var1"),
         yaxis = list(title = "var2")) 

```