---
title: "Network data availability"
author: "Anna Krystalli"
date: "14 December 2015"
output: html_document
---


```{r, echo=F, message=FALSE, warning= F}

codewd <- "~/Documents/worflows/bird_trait_networks/"
datawd <- "~/"

output.folder <- "/Users/Anna/Google Drive/bird trait networks/outputs/"
input.folder <- "/Users/Anna/Google Drive/bird trait networks/inputs/data/"


require(dplyr)
require(plotly)
require(knitr)
require(RColorBrewer)

D0 <- read.csv(file = paste(input.folder, "csv/","D0.csv", sep = "")
               ,fileEncoding = "mac") 

numerise <- function(x){if(all(grepl('^[0-9.]+$', x))) as.numeric(x) else x}

makeTransparent<-function(someColor, alpha=100)
{
  newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                              blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
}

#source("~/Documents/worflows/bird_trait_networks/Setup.R")

metadata <- read.csv(paste(input.folder, "metadata/","metadata.csv", sep = ""), 
                     stringsAsFactors = F, fileEncoding = "mac") %>% 
  apply(2, FUN = trimws) %>% data.frame(stringsAsFactors = F)


vnames <- read.csv(paste(input.folder, "metadata/","vnames.csv", sep = ""), 
                   stringsAsFactors = F, fileEncoding = "mac") %>% 
  apply(2, FUN = trimws) %>% data.frame(stringsAsFactors = F)
vnames[vnames == ""] <- NA

#__________________________________________________________________
### Data availability across **`D0`** bivariate combinations
```

I've calculated counts of data available for each possible bivariate combination of variables in the dataset. Below is the distribution of counts across all bivariate combination.



```{r, echo=F, message=FALSE, warning= F}

load(file = paste(input.folder, "r files/n.comb.Rdata", sep = ""))

# make grid with unique combinations of variables
var.grid <- expand.grid(unique(D0$var), unique(D0$var))


# make matrix
m <- matrix(NA, ncol = length(unique(D0$var)), nrow = length(unique(D0$var)),
dimnames = list(unique(D0$var), unique(D0$var)))

m[cbind(var.grid[,1], var.grid[,2])] <- n.comb

plot_ly(x = n.comb, type = "histogram") %>%
  layout(title = "distribution of bivariate datapoint availability",
         xaxis = list(title = "number of datapoints available per bivariate combination"),
         yaxis = list(title = "frequency"))

cat <- metadata$cat[match(dimnames(m)[[1]], metadata$master.vname)]
text <- cat[order(cat)]

m <- m[order(cat),order(cat)]
```


I've also plotted interactively (*ie hover over the matrix for details of each cell*) the matrix of bivariate counts, sorted by variable category (ie they are grouped and ordered by what type of variable they are (ie *behavioural, ecological etc*); let me know if you want me to sort the matrix some other way):
  
  
```{r, echo=F, message=FALSE, warning= F}
col <- brewer.pal(5, "YlGnBu")

plot_ly(z = m, x = unique(D0$var), y = unique(D0$var), type = "heatmap", 
        #color = col, 
        name = "bivariate datapoint availability",
        colors = colorRamp(col, bias = 2),
        hoverinfo = "z+x+y+text", text = text) %>%
  
  layout(margin = list(l = 150,
                       r = 80,
                       b = 120,
                       t = 60),
         title = "matrix of bivariate datapoint availability",
         xaxis = list(title = "var1"),
         yaxis = list(title = "var2")) 

```