---
title: "Categorical Phylocors"
author: 
date: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
wkf = "phylonetworker"
param = "phylonetworker.R"
file_setup_path <- "~/Documents/workflows/bird_trait_networks/file_setup.R"
source(file_setup_path)
source(paste0(script.folder, "params/", param))

makeTransparent <- function(someColor, alpha=100){
  newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                              blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
}
```


```{r}
load(file = paste(output.folder, "data/phylocors/", an.ID,"_phylocor_mn", 
                         min.n, if(log){"_log"},"_cc.Rdata", sep = ""))

```

```{r, echo=FALSE}
all <- NULL
obs <- NULL
diff.max <- NULL
diff.min <- NULL

valid.ids <- (1:length(cc.res))[sapply(cc.res, FUN = length) != 1]
for(i in valid.ids){
  x <- cc.res[[i]]
    
  all <-  c(all, 
            max(sapply(x$phy.GKtau, FUN = function(x){
              mean(replace(x, is.infinite(x),NA),na.rm = T)})))

  obs <- c(obs, max(x$GKtau[,5:6]))
  max.id <- which(x$GKtau[,5:6] == max(x$GKtau[,5:6]))[1]
  min.id <-setdiff(1:2, max.id)
  diff.max <- c(diff.max, x$GKtau[,5:6][max.id] - mean(replace(x$phy.GKtau[,max.id],
                                                            is.infinite(x$phy.GKtau[,max.id]),
                                                            NA), 
                                                       na.rm = T))
  diff.min <- c(diff.min, x$GKtau[,5:6][min.id] -
                  mean(replace(x$phy.GKtau[,min.id],
                               is.infinite(x$phy.GKtau[,min.id]), 
                               NA),
                       na.rm = T))
} 

xlims <- range(unlist(c(diff.min, diff.max)))

```

## Density distribution of **mean** simulated trait pair GK $τ$ arising purely by phylogeny **across all trait pairs** (red) vs distribution of observed GK $τ$ (black).

Red line indicates mean GK $τ$ across all simulations. Black line, mean of all observed GK $τ$. For each pair, the mean GK $τ$ of each trait on the other was calculated across all simulations and the maximum for each pair retained (remember the asymmetry). Similarly, for the observed GK $τ$, the maximum was retained for each pair.

```{r, echo=FALSE}
plot(density(all), xlim = range(c(all, obs)))
polygon(density(all), col=makeTransparent("red", 60), border="red")
polygon(density(obs), col=makeTransparent("black", 20), border="black")
abline(v = mean(all), col = "red", lwd = 2)
abline(v = mean(obs), lwd = 2)
abline(v = 0, lty = 2, col = "blue")

```

### Density distribution of difference between **max observed GK $τ$** and **mean simulated trait pair GK $τ$** arising purely by phylogeny **across all trait pairs** (*remember GK $τ$ asymmetric hence maximum selected*).

So value of 0 means no difference between phylogenetically simulated and observed GK $τ$, positive values indicate observed GK $τ$ being larger than phylogenetically simulated and negative values the opposite.

```{r}

plot(density(unlist(diff.max)))
  polygon(density(unlist(diff.max)), col=makeTransparent("red", 60), border="red")
  abline(v = 0, lty = 2)
```

### Density distribution of difference between **min observed GK $τ$** and **mean simulated trait pair GK $τ$** arising purely by phylogeny **across all trait pairs**

```{r}
plot(density(unlist(diff.min)))
  polygon(density(unlist(diff.min)), col=makeTransparent("blue", 60), border="blue")
  abline(v = 0, lty = 2)

```


# Individual pair plots

Red areas indicate the density distribution of **simulated GK $τ$ based purely on phylogeny** for trait with the highest **observed** GK $τ$ on the other (remember the asymmetry). Blue indicates the lower observed GK $τ$. Solid lines indicate the **observed GK $τ$** while dashed lines indicate the mean of the **phylogenetically simulated GK $τ$s** (n = 500).

The right panels show the frequency among **phylogenetically simulated GK $τ$s** in which the effect of each trait on the other was greatest.

```{r, fig.width=10, results='asis'}

par(mfrow = c(1,2))
#
for(i in valid.ids[order(unlist(diff.max), decreasing = T)]){
    x <- cc.res[[i]]
    if(length(x) == 1){
      cat("### ", x)
      next}

#split.screen(figs = matrix(c(0,0.65,0,1, 0.65,1, 0,1), nrow = 2, byrow=TRUE))
#screen(1)


  x$phy.GKtau[,1] <- replace(x$phy.GKtau[,1], is.infinite(x$phy.GKtau[,1]), NA)
  x$phy.GKtau[,2] <- replace(x$phy.GKtau[,2], is.infinite(x$phy.GKtau[,1]), NA)
  

  max.trait <- unlist(apply(x$phy.GKtau, 1, 
                            FUN = function(x){
                              which(x == max(x, na.rm = T), arr.ind = T)
                              }))
  
  max.id <- which(x$GKtau[,5:6] == max(x$GKtau[,5:6]))[1]
  min.id <-setdiff(1:2, max.id)
  
  d1 <- density(x$phy.GKtau[,max.id], na.rm = T)
  d2 <- density(x$phy.GKtau[,min.id], na.rm = T)
  plot(d1, main = x$pair,xlim = range(c(d1$x, d2$x, x$GKtau[,5:6][max.id])),
       ylim = range(c(d1$y, d2$y)))
  polygon(d1, col=makeTransparent("red", 60), border="red")
  polygon(d2, col=makeTransparent("blue", 40), border="blue")
  abline(v = mean(x$phy.GKtau[,max.id], na.rm = T), col = "red", lwd = 2, lty = 2)
  abline(v = mean(x$phy.GKtau[,min.id], na.rm = T), col = "blue", lwd = 1, lty = 2)

  abline(v = x$GKtau[,5:6][max.id], lwd = 3, col = "red")
  abline(v = x$GKtau[,5:6][min.id], lwd = 2, col = "blue")
  
  
  
legend("topright", paste0("x: ", x$pair[c(max.id, min.id)]), bty = "n", fill = c("red", "blue"), cex = 0.5)
#screen(2)
barplot(table(max.trait), names.arg = x$pair[as.numeric(names(table(max.trait)))], main = "freq of trait GK τ being highest")
}



```

