---
output: html_document
---



```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=9}

# SETTINGS
require(knitr)
require(igraph)
library(DT)

output.folder <- "/Users/Anna/Google Drive/bird trait networks/outputs/"
input.folder <- "/Users/Anna/Google Drive/bird trait networks/inputs/data/"

an.ID <- "100spp"
log <- T
min.n <- 10
cutoff <- 0.35

# FILES
load(file = paste(output.folder, "data/networks/", an.ID,"_net_mn", min.n, 
                       if(log){"_log"},".RData", sep = ""))

res <- read.csv(paste(output.folder, "data/phylocors/", an.ID,"_phylocor_mn", 
                     min.n, if(log){"_log"},".csv", sep = ""), stringsAsFactors = F)

metadata <- read.csv(paste(input.folder, "metadata/","metadata.csv", sep = ""), 
                     stringsAsFactors = F, fileEncoding = "mac") %>% 
  apply(2, FUN = trimws) %>% data.frame(stringsAsFactors = F)

### FUNCTION


makeTransparent<-function(someColor, alpha=100)
{
  newColor <- col2rgb(someColor)
  newColor <- rbind(newColor, alpha)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                              blue=curcoldata[3], alpha=curcoldata[4],
                                              maxColorValue=255)})
}

```

# **Correlation network report**

##### **Analysis ID:   ** `r an.ID` 
##### **log:   ** `r log`
##### **min.n:   ** `r min.n`
##### **cutoff:   ** `r cutoff`

The report visualises the results of network analysis of phylogenetically corrected correlations $(r_p)$ between numeric traits. $(r_p)$ values where extracted from `pgls` models which corrected for phylogenetic bias. Models were attempted only for trait combinations for which data availability exceeded **`r min.n`** data points. Traits where connected if the absolute $r_p$ values between them exceeded a cut-off point of: **`r cutoff`**. The associated script can be found [here](https://github.com/annakrystalli/bird_trait_networks/blob/master/PhyloNetworker.R).



```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=9}

################################################################################################
# PREPARE REPORT DATA


edges <- res[abs(res$phylocor) > cutoff & !is.na(res$phylocor),]

g.n <- as.vector(apply(edges[,c("var1", "var2")],1,FUN = unlist))
g.names <- unique(g.n)

vertices <- data.frame(code = 1:dim(net[[1]])[1], net[[1]][match(g.names, net[[1]]$name),]) 



g.code <- match(g.n, g.names)
v.code.n <- match(g.names, vertices$name)
v.code.n <- match(vertices$name, g.names)



# node colours
all.role.levels <- levels(vertices$role)
role.levels <- all.role.levels[which(all.role.levels %in% vertices$role)]

mod.levels <- sort(unique(vertices$module))

# role.codes <- 
role.cols <- RColorBrewer::brewer.pal(length(role.levels), "BrBG")
mod.cols <- RColorBrewer::brewer.pal(length(mod.levels), "Set2")
names(mod.cols) <- mod.levels
role.size <- rev(1/1:length(all.role.levels))[all.role.levels %in% role.levels]
names(role.size) <- role.levels


vertex.cols <- mod.cols[as.character(vertices$module)]
n.size <- role.size[as.character(vertices$role)]

# edge colours
edge.cols <- rep("black", length(edges$phylocor))
edge.cols[edges$phylocor >= 0.1] <- "red"
edge.cols[edges$phylocor < -0.1] <- "blue"


require(igraph)

G <- graph(g.code, directed = FALSE)

# Assign attributes to the graph
G$name    <- "correlation network of numeric bird traits"

# Assign attributes to the graph's vertices
# V(G)$name  <- 1:length
V(G)$color <- vertex.cols
E(G)$edge.color <- edge.cols


# Assign attributes to the edges
E(G)$weight <- edges$phylocor



# re-order vertices into data
meta.id <- match(gsub("log_", "", g.names), metadata$code)
data <- data.frame(code = 1:length(g.names), name = g.names,
                   descr = metadata$descr[meta.id], module = vertices$module, 
                   role = vertices$role, connectivity = vertices$connectivity,
                   participation = vertices$participation,
                metadata[meta.id,][,c("cat", "units")])

# phylocors

g.n.mod <- data$module[match(g.n, data$name)]
names(g.n.mod) <- g.n

intra.id <- tapply(g.n, rep(1:(length(g.n)/2), each = 2), 
                   FUN = function(x, g.n.mod){
                     g.n.mod[unlist(x)][1] == g.n.mod[unlist(x)][2]}, 
                   g.n.mod = g.n.mod)

intra.e <-edges[intra.id,]
intra.e$module <- g.n.mod[rep(intra.id, each = 2)][seq(2, 2*sum(intra.id), by = 2)]


```

<br>

### **network characteristics**

##### **no. modules:** `r length(mod.levels)`
##### **no. vertices:** `r nrow(vertices)`
##### **no. edges:** `r nrow(edges)`
##### **mean participation:** `r format(mean(vertices$participation), digits = 2)`
##### **mean connectivity:** `r format(mean(vertices$connectivity), digits = 3, nsmall = 2)`


## **modules**

#### vertex characteristics
```{r, results="asis", echo=F, warning=F, message=F, fig.width=4, fig.height=3.5}


par(mai = c(4,2,2,4), las = 1, mar = c(4,4,2,1), mgp = c(2.2, 1, 0))

boxplot(connectivity ~ module, data = data, xlab = "modules",
        main = "a) distribution of vertex connectivity between modules", cex.main = 0.75,
        varwidth = T, col = mod.cols[as.character(sort(unique(data$module)))], notch = T,
        ylab = "connectivity",
        cex.lab = 0.7, cex.axis = 0.7, bty = "n")

boxplot(participation ~ module, data = data,  xlab = "modules", outline = F,
        main = "b) distribution of vertex participation between modules", cex.main = 0.75,
        varwidth = T, col = mod.cols[as.character(sort(unique(data$module)))], notch = T,
        ylab = "participation",
        cex.lab = 0.7, cex.axis = 0.7)
```

#### edge characteristics

```{r, results="asis", echo=F, warning=F, message=F, fig.width=4, fig.height=3.5}

boxplot(abs(phylocor) ~ module, data = intra.e, xlab = "modules", outline = T,
        main = "c) distribution of abs(intra-modular) phylocor between modules", cex.main = 0.75,
        varwidth = T, col = mod.cols[as.character(sort(unique(data$module)))], notch = T,
        ylab = "abs(phylocorrected correlation)",
        cex.lab = 0.7, cex.axis = 0.7, ylim = c(0, 1))

boxplot(phylocor ~ module, data = intra.e, xlab = "modules", outline = T,
        main = "d) distribution of intra-modular phylocor between modules", cex.main = 0.75,
        varwidth = T, col = mod.cols[as.character(sort(unique(data$module)))], notch = T,
        ylab = "phylocorrected correlation",
        cex.lab = 0.7, cex.axis = 0.7, ylim = c(-1, 1))

```

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=9}
par(mai = c(0,0.5,0.5,0))

set.seed(1)
plot(G, layout = layout.fruchterman.reingold, 
     main = G$name,
     #vertex.label = V(G)$name,
     vertex.size = 30*n.size,
     vertex.color= vertex.cols,
     vertex.frame.color= vertex.cols,
     vertex.label.color = "black",
     vertex.label.family = "sans",
     edge.width=E(G)$weight,
     vertex.label.cex=2.2 * n.size,
     edge.color= makeTransparent(edge.cols, 70))


legend("bottomleft", legend = names(role.size), 
       col = "black", pch = 21, pt.bg = "black", title = "roles",
       bty = "n", pt.cex = 4 * role.size)

legend("bottomright", legend = names(mod.cols), 
       col = mod.cols, pch = 21, pt.bg = mod.cols, title = "modules",
       bty = "n", pt.cex = c(1))



dt.tab <- data
dt.tab <- dt.tab[order(dt.tab$participation, decreasing = T),]
dt.tab$connectivity <- format(dt.tab$connectivity, digits = 4)
dt.tab$participation <- format(dt.tab$participation, digits = 3, n.small = 2,
                                          scientific = 9999, trim = F)



DT::datatable(dt.tab, rownames = F, filter = list(position = 'top'))

rm(dt.tab)

dev.off()

```

<br>
 
```{r, message=FALSE, warning=FALSE, echo=FALSE, include = F}

out <- NULL
for(mod in as.character(sort(mod.levels))){
  
  out <- c(out, 
           knit_child("~/Documents/workflows/bird_trait_networks/reports/results/module_plots_child.Rmd", 
                      envir = parent.frame()))
  }
```
```{r echo=FALSE, results="asis"}

cat(paste(out, collapse = "/n"))

```

