---
title: "Network_viz"
date: "Last rendered: `r format(Sys.time(), '%d %b %Y %H:%M:%S')`"
output: 
  html_document:
    theme: paper
params:
  wkf: phylonetworker
  param: phylonetworker.R
  file_setup_path: ~/Documents/workflows/bird_trait_networks/file_setup.R
  an.ID: "param"
  edge_det: "param"
  node_colour: "param"
---
```{r nviz-setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
options(stringsAsFactors=FALSE, scipen = 0)

attach(params)
param
source(params$file_setup_path)
source(paste(script.folder, "params/", params$param, sep = ""))
source(paste0(script.folder,"project_ui.R"))
```

```{r, echo=FALSE}
if(params$an.ID != "param"){an.ID <- params$an.ID}
if(params$edge_det != "param"){edge_det <- params$edge_det}
if(params$node_colour != "param"){node_colour <- params$node_colour}
```

## Analysis parameters
- **Analysis method:** `r wkf`
- **edge determination method:** `r edge_det`
- **species subset:** `r an.ID`
- **minimum $n$ for edge determination:** `r min.n`
- **log numeric variables?:** `r log`
- **edge correlation cutoff value:** `r cutoff`
- **Minimum category $n$ (below which aggregated or removed):** `r min.cat`




## Analysis diagnostics


```{r, echo = F}
res <- read.csv(paste(output.folder, "data/phylocors/", an.ID,"_phylocor_mn", 
                      min.n, if(alco){"alco_"}, if(log){"_log"},"_allDT.csv", sep = ""))

net <- read.csv(paste(output.folder, "data/networks/", an.ID,
                    "_net_mn", min.n, 
                    "_", edge_det,
                    if(alco){"alco_"},
                    if(log){"_log"},".csv", sep = ""))

net_stats <- read.csv(paste(output.folder, "data/networks/", an.ID,
                    "_net_mn", min.n, 
                    "_", edge_det,
                    if(alco){"alco_"},
                    if(log){"_log"},"_net-stats.csv", sep = ""))



```

- **Number of traits (potential nodes):** `r net_stats$nvar`
- **Number of pairs with enough data to determine edges (ie $n>n_{min}$):** `r net_stats$pairs` (`r round((net_stats$pairs/net_stats$tpairs * 100), digits = 0)`%)
- **Number of edge determinations failed to converge:** `r net_stats$errors` (`r round((net_stats$errors/net_stats$pairs * 100), digits = 0)`%)

### Results

- **Network modularity:** `r net$modularity[1]`
- **Network realised modularity (eq. 2):** `r net_stats$q_r2`
- **Number of nodes in final network:** `r net_stats$nodes`
- **Number of modules:** `r length(unique(net$module))`
- **Number of edges determinationed by `phylocor` > `cut.off`:** `r net_stats$edges.phy`
- **Number of edges determinationed by `p` < 0.05:** `r net_stats$edges.p`

```{r, echo=FALSE}
if(edge_det == "phylocor"){
res <- res[abs(res$phylocor) > cutoff & !is.na(res$phylocor),]
}else{
 res <- res[res[,"p"] < 0.05 & !is.na(res[,"p"]),] 
}
```


# Network

```{r, eval = T}
shapes <- setNames(c("dot","circle","ellipse", "triangle", "square",
           "database","diamond"),
         c("Ultra peripheral", "Peripheral", "Connector", "Kinless",
           "Peripheral Hub", "Connector Hub", "Kinless Hub"))[unique(net$role)]

addNodes <- data.frame(label = names(shapes), shape = shapes, size = 2)

net <- net[order(net$name),]

nodes <- data.frame(id = 1:nrow(net),
                    label = net$name,
                    title = net$name,
                    group = net[,node_colour],
                    value = plotrix::rescale(net$connectivity, c(8.5,18)),
                    shape = shapes[net$role],
                    net)
#nodes$colour.background

links <- res[, c("var1","var2","n")]
links$width <- abs(res$phylocor) * 5
links$from <- nodes$id[match(links$var1, nodes$title)]
links$to <- nodes$id[match(links$var2, nodes$title)]
links$colour <- "red"
links$colour[res$phylocor < 0] <- "blue"
links$colour[res$pair.type == "cc"] <- "grey"

visNetwork(nodes, links, height = "600px", width = "100%") %>% 
   #visEdges(color = list(color = links$colour,
                         #highlights = links$colour)) %>%
  visPhysics(maxVelocity = 25) %>% 
    
  #  solver = "barnesHut",
   #          barnesHut = list(gravitationalConstant = -1750,
    #                          springConstant = 0.1)) %>%
  #visEdges(value = "weight", arrows = "to") %>% 
  visOptions(highlightNearest = TRUE, selectedBy = "module", nodesIdSelection = T) %>% 
  visLegend(addNodes = addNodes)


```



## Modules

- **total number of modules:** `r length(unique(net$module))`
- **distribution of nodes across nodules:**

### module size (number of nodes)
```{r, fig.width = 8, fig.height = 4, eval = T}
ggplot(data = net, aes(x = as.factor(module))) +
  geom_bar(aes(fill = cat))
```

### connectivity
```{r, fig.width = 8, fig.height = 3.2, eval = T}
ggplot(data = net, aes(x = as.factor(module), y = connectivity, fill = as.factor(module))) +
  geom_boxplot()
```

### participation
```{r, fig.width = 8, fig.height = 3.2, eval = T}
ggplot(data = net, aes(x = as.factor(module), y = participation, fill = as.factor(module))) + geom_boxplot()
```

### proportion of links tested (per trait) which were significant
```{r, fig.width = 8, fig.height = 3.2, eval = T}
ggplot(data = net, aes(x = as.factor(module), y = p.prop, fill = as.factor(module))) + geom_boxplot()
```

### trait phylogenetic signal
```{r, fig.width = 8, fig.height = 3.2, eval = T}
ggplot(data = net, aes(x = as.factor(module), y = phylosig, fill = as.factor(module))) + geom_boxplot()
```

### median phylocor
```{r, fig.width = 8, fig.height = 3.2, eval = T}
ggplot(data = net, aes(x = as.factor(module), y = phycor_median, fill = as.factor(module))) + geom_boxplot()
```

### taxonomic distinctness
```{r, fig.width = 8, fig.height = 3.2, eval = T}
ggplot(data = net, aes(x = as.factor(module), y = Dplus, fill = as.factor(module))) + geom_boxplot()
```

### pairs_tested
```{r, fig.width = 8, fig.height = 3.2, eval = T}
ggplot(data = net, aes(x = as.factor(module), y = pairs_tested, fill = as.factor(module))) + geom_boxplot()
```

***
# Data
## Node level and network data.

`rnetcarto::netcarto()` output. For details on the [package](https://cran.r-project.org/web/packages/rnetcarto/vignettes/getting-started.html) and [theory](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2151742/). Background to calculation of realised modularity in this [paper](https://f1000research.com/articles/2-130/v3). Equation for $Q'_r$ used.

```{r ktab, eval = F}
DT::datatable(format(net, digits = 2, scientific = F), filter = list(position = 'top'), rownames = F)
```



- **name:** trait name
- **module:** module membership
- **connectivity:** ($k_i$) within-module degree *z-score*. Measures how ‘well connected’ node is to other nodes in it's module.
- **participation:** ($P_i$) close to one if a node's links are uniformly distributed among all the modules and zero if all its links are within its own module.
- **role**: role of trait in the network. See below for definitions.
- **descr:** full description of trait
- **cat:** trait category
- **type:** trait type
- **scores:** scores used to code factors
- **levels:** levels associated with factor codes
- **EDplus:** expected taxonomic distinctness given sample size
- **Dplus:** species sample estimate of taxonomic distinctness.
- **sd.Dplus:** standard deviation of taxonomic distinctness estimate.
- **phylosig:** trait phylogenetic signal
- **p.prop:** proportion of links involving trait tested which where significant.
- **pairs_tested:** number of pairs involving trait tested.
- **phylocor_mean:** mean phylocorrected correlation across pairs involving trait tested.
- **phylocor_median:** median phylocorrected correlation across pairs involving trait tested.
- **phylocor_max:** max phylocorrected correlation across pairs involving trait tested.


```{r ntab, eval = T}

DT::datatable(format(net, digits = 1, scientific = F, signif = 4), filter = list(position = 'top'), rownames = F)

```

### Network roles
#### Non-hubs

- **Ultra-peripheral nodes (role R1).** If a node has all its links within its module ($P ≈ 0$).

- **Peripheral nodes (role R2).** If a node has at least 60% its links within the module, then for $k < 4$ it follows that $P < 0.625$.

- **Non-hub connectors (role R3).** If a node with $k < 4$ has half of its links (or at least two links, whichever is larger) within the module, then it follows that $P < 0.8$. Thus, a plausible region for non-hub connectors is $0.62 < P < 0.8$.

- **Non-hub kinless nodes (role R4).** If a node has fewer than 35% of its links within the module, it implies that $P > 0.8$. We surmise that such nodes cannot be clearly assigned to a single module. We thus classify them as kinless nodes. We will demonstrate later that non-hub kinless nodes are found in most network growth models, but not in real-world networks.

#### Hubs

- **Provincial hubs (role R5).** If a node with a large degree, $k » 1$, has at least 5/6 of its links within the module, then it follows that $P = 1 − (5/6)2 − (k/6)(1/k2) = 0.31 − 1/(6k) ≈ 0.30$.

- **Connector hubs (role R6).** If a node with a large degree has at least half of its links within the module, then it follows that $P = 1 − 1/4 − (k/2)(1/k2) = 0.75 − 1/(2k)$. Since $k » 1$, $P < 0.75$ for such nodes.

- **Kinless hubs (role R7).** If a hub has fewer than half its links within the module, i.e., $P > 0.75$, then we surmise that it may not be clearly associated with a single module. We then classify it as a kinless hub. We will demonstrate later that hubs in most network growth models are actually kinless hubs.

***
<br>

## Edges data 

Results of phylogentically corrected trait correlation determination. See analysis documentation for details:

- **n:** number of data used in edge estimation
- **Dplus:** species subsample estimate of taxonomic distinctness.
- **sd.Dplus:** standard deviation of taxonomic distinctness estimate.
- **EDplus:** expected taxonomic distinctness given subsample size
- **physig.var1:** phylogenetic signal in var1
- **physig.var2:** phylogenetic signal in var2
- **cor:** standard correlation
- **phylocor:** phylogenetically corrected correlation
- **lambda:** $λ$
- **p:** $p$
- **l.ci.l:** $λ$ lower confidence interval
- **l.ci.u:** $λ$ upper confidence interval
- **error:** error message for failed determnations
- **pair.type:** pair variable types: 
    - `"nn"` numeric vs numeric
    - `"nc"` numeric vs categorical
    - `"cc"` categorical vs categorical


```{r etab, eval = T}

DT::datatable(format(res, digits = 1, scientific = F, signif = 4), filter = list(position = 'top'), rownames = F)

```




