---
title: "Network_viz"
date: "Last rendered: `r format(Sys.time(), '%d %b %Y %H:%M:%S')`"
output: 
  html_document:
    theme: paper
params:
  wkf: phylonetworker
  param: phylonetworker.R
  file_setup_path: ~/Documents/workflows/bird_trait_networks/file_setup.R
---
```{r nviz-setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
options(stringsAsFactors=FALSE, scipen = 0)
require(knitr)
library(tidyr)
library(plotrix)
require(DT)
require(visNetwork)

attach(params)
param
source(file_setup_path)
source(paste(script.folder, "params/", param, sep = ""))

if(log){
  metadata <- read.csv(paste(input.folder, "metadata/","metadata.csv", sep = ""),
                       stringsAsFactors = F, fileEncoding = "mac", 
                       na.strings = c("","NA", " "), strip.white = TRUE, 
                       blank.lines.skip = TRUE)
  log.vars <- metadata$code[as.logical(metadata$log)]
}


```


## Analysis parameters
- **edge determination method:** `r wkf`
- **species subset:** `r an.ID`
- **minimum $n$ for edge determination:** `r min.n`
- **log numeric variables?:** `r log`
- **edge correlation cutoff value:** `r cutoff`
- **Minimum category $n$ (below which aggregated or removed):** `r min.cat`




## Analysis diagnostics


```{r, echo = F}
res <- read.csv(paste(output.folder, "data/phylocors/", an.ID,"_phylocor_mn", 
                      min.n, if(log){"_log"},"_allDT.csv", sep = ""))

net <- read.csv(paste(output.folder, "data/networks/", an.ID,"_net_mn", min.n, 
                      if(log){"_log"},".csv", sep = ""))

nvar <- length(unique(unlist(res[,1:2])))
tpairs <- choose(nvar,2) 
errors <- sum(!is.na(res$error)) - length(grep("ERROR: singular category:", res$error))
pairs <- nrow(res) - length(grep("ERROR: singular category:", res$error))
edges <- dim(res[abs(res$phylocor) > cutoff & !is.na(res$phylocor),])[1]
nodes <- nrow(net)
```

- **Number of traits (potential nodes):** `r nvar`
- **Number of pairs with enough data to determine edges (ie $n \ge n_{min}$):** `r pairs` (`r round((pairs/tpairs * 100), digits = 0)`%)
- **Number of edge determinations failed to converge:** `r errors` (`r round((errors/pairs * 100), digits = 0)`%)

### Results

- **Number of nodes in final network:** `r nodes`
- **Number of modules:** `r length(unique(net$module))`

```{r, echo=FALSE}
res <- res[abs(res$phylocor) > cutoff & !is.na(res$phylocor),]
```


### network

```{r}
shapes <- setNames(c("dot","circle","ellipse", "triangle", "square",
           "database","diamond"),
         c("Ultra peripheral", "Peripheral", "Connector", "Kinless",
           "Peripheral Hub", "Connector Hub", "Kinless Hub"))[unique(net$role)]

addNodes <- data.frame(label = names(shapes), shape = shapes, size = 2)

net <- net[order(net$name),]

nodes <- data.frame(id = 1:nrow(net),
                    label = net$name,
                    title = net$name,
                    group = net$module,
                    value = plotrix::rescale(net$connectivity, c(8.5,18)),
                    shape = shapes[net$role],
                    net)
#nodes$colour.background

links <- res[, c("var1","var2","n")]
links$width <- res$phylocor
links$from <- nodes$id[match(links$var1, nodes$title)]
links$to <- nodes$id[match(links$var2, nodes$title)]


visNetwork(nodes, links, height = "600px", width = "100%") %>% 
  visPhysics(maxVelocity = 25) %>% 
    
  #  solver = "barnesHut",
   #          barnesHut = list(gravitationalConstant = -1750,
    #                          springConstant = 0.1)) %>%
  #visEdges(value = "weight", arrows = "to") %>% 
  visOptions(highlightNearest = TRUE, selectedBy ="module", nodesIdSelection = T) %>% 
  visLegend(addNodes = addNodes)


```



#### modules

- **total number of modules:** `r length(unique(net$module))`
- **distribution of nodes across nodules:**

```{r, fig.width = 3, fig.height = 2}
barplot(table(net$module), col = "darkgreen")
```

### Node level and network data.

`rnetcarto::netcarto()` output. For details on the [package](https://cran.r-project.org/web/packages/rnetcarto/vignettes/getting-started.html) and [theory](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2151742/).

```{r ktab}
DT::datatable(format(net, digits = 2, scientific = F), filter = list(position = 'top'), rownames = F)
```

***

### Edges data 

Results of phylogentically corrected trait correlation determination. See annalysis documentation for details:

- **n:** number of data used in edge estimation
- **Dplus:** species subsample estimate of taxonomic distinctness.
- **sd.Dplus:** standard deviation of taxonomic distinctness estimate.
- **EDplus:** expected taxonomic distinctness given subsample size
- **physig.var1:** phylogenetic signal in var1
- **physig.var2:** phylogenetic signal in var2
- **cor:** standard correlation
- **phylocor:** phylogenetically corrected correlation
- **lambda:** $λ$
- **p:** $p$
- **l.ci.l:** $λ$ lower confidence interval
- **l.ci.u:** $λ$ upper confidence interval
- **error:** error message for failed determnations
- **pair.type:** pair variable types: 
    - `"nn"` numeric vs numeric
    - `"nc"` numeric vs categorical
    - `"cc"` categorical vs categorical


```{r ntab}

DT::datatable(format(res, digits = 1, scientific = F, signif = 4), filter = list(position = 'top'), rownames = F)

```


