## PhyloCors Workflow:


```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(caper)
library(geiger)
library(PHYLOGR)
```

**`tree`:** a phylogenetic tree

```{r}
load("/Users/Anna/Google Drive/bird trait networks/inputs/data/tree/example tree.RData")
str(tree)

```

<br>

***

#### First part of workflow sets up two simulated traits (not required)

**`rTraitCont {ape}`:** simulates the evolution of a continuous character along a phylogeny. The calculation is done recursively from the root. (*Paradis, 2012, pp. 232 and 324 for intro*). 

- Default `model` is **`"BM"`** : a Browian motion model is used. If the arguments `sigma` has more than one value, its length must be equal to the the branches of the tree. This allows to specify a model with variable rates of evolution. You must be careful that branch numbering is done with the tree in “pruningwise” order: to see the order of the branches you can use: `tr <- reorder(tr, "p"); plor(tr); edgelabels()`. The arguments `alpha` and `theta` are ignored.
- `sigma` = 0.1, 
- `ancestor` = `FALSE`, 
- `root.value` = 0

**`rescale {geiger}`:** applies various transformation to the branches of a phylogenetic tree.

- **`lambda`:** is one of the Pagel (1999) models that fits the extent to which the phylogeny  predicts covariance among trait values for species. The model effectively transforms the tree as follows: values of **lambda near 0** cause the **phylogeny to become more star-like**, and a lambda value of **1** recovers the **BM model** (Brownian motion model: random walk with variance `sigsq`. The parameter used for transformation is `lambda`.

**Q: Why are we rescaling phylogeny with a BM model?**

```{r}
set.seed(1)
x<- scale(rTraitCont(rescale(tree, model = "lambda", 1)))
set.seed(2)
y<- scale(rTraitCont(rescale(tree, model = "lambda", 1)))

head(x)
head(y)

```


***

#### Second part: calculates the **Phylogenetically Independent Contrasts**

**`pic {ape}`:** Compute the phylogenetically independent contrasts using the method described by *Felsenstein* (1985).

**`pic(x, phy, scaled = TRUE, var.contrasts = FALSE, rescaled.tree = FALSE)`**

- **`x`:** a numeric vector
- **`phy`:** an object of class "phylo"
- **`scaled`:** logical, indicates whether the contrasts should be scaled with their expected variances. 
- **`var.contrasts`:** logical, indicates whether the expected variances of the contrasts should be returned
- **`rescaled.tree`:** logical, if TRUE the rescaled tree is returned together with the main results.
```{r} 
x.pic<- pic(x, tree)
head(x.pic)
y.pic<- pic(y, tree)
head(y.pic)
```

**`cor.origin {PHYLOGR}`:** Return the correlation through the origin of two vectors. ie. The correlation of x and y, from a model without intercept (i.e., forcing the line through the origin). Generally used for indepdendent contrasts.

```{r}
cor.origin(x.pic,y.pic) # this will equal the cor.xy below

```


***

**`vcv {ape}`:** estimating the evolutionary or phylogenetic variance-covariance matrix.

**`vcv(phy, model = "Brownian", corr = FALSE, ...)`**

+   This function computes the **expected variances** and **covariances** of a **continuous trait** ***assuming it evolves under a given model***. Returns a numeric matrix with the names of the tips as colnames and rownames.

    - **`phy`:an object of the correct class ('phylo' or 'corPhyl')
    - **`model`:** a character giving the model used to compute the variances and covariances; only "`Brownian`" is available (for other models, a correlation structure may be used).
    - **`corr`:** a logical indicating whether the correlation matrix should be returned (TRUE); by default the variance-covariance matrix is returned (FALSE).


**`solve {base}`:** solves a system of equations. ie **`a %*% x = b`**. if **`b`**, it is taken to be an identity matrix and `solve` will return the inverse of **`a`**.

**`%*%`:** perform matrix multiplication

```{r, eval=FALSE}
invC <-solve(vcv.phylo(tree))
```

So the above command computes the expected covariances and inverts them so instead of the strength of correlation between traits it **actually returns a measure of trait dissimilarity??**

Inverting the matrix by the way is by far the most time consuming step

```{r, echo = FALSE}
load("/Users/Anna/Google Drive/bird trait networks/inputs/data/tree/example invC.RData")
```

<br>

In the next step the matrix is multiplied by the trait values for each variable
```{r}
mean.x <-colSums(invC%*%x)/sum(invC)
mean.y <-colSums(invC%*%y)/sum(invC)

```

In var x, first the deviation of each variable from mean.x in multiplied by 
```{r}
nsps <- length(tree$tip.label)
vector.ones<-as.matrix(rep(1,nsps))

var.x <-t(x-mean.x) %*% invC%*%(x-mean.x)/(nsps-1)
var.y <-t(y-mean.x) %*% invC%*%(y-mean.x)/(nsps-1)
cor.xy <-(t(x-mean.x) %*% invC%*%(y-mean.x)/(nsps-1))/sqrt(var.x*var.y)
cor.xy




```
